---
title: "Data EXtraction report for Occupancy rate KPI"
output: html_notebook
---

```{r}
library(dplyr)
data_folder <- paste(getwd(), '/../data/raw/', sep='')
stage_folder <- paste(getwd(), '/../data/stage/', sep='')
```

Reading Occupation dataset
```{r}
dfoccupation<-read.csv(paste(data_folder, "ocupacao_hospitalar_2011-2016-utf8.csv", sep=''), sep=';', header=T)
dfsector<-read.csv(paste(data_folder, "setor_atendimento_2017-08-04.csv", sep=''), sep=',', header=T)
```

Converting column names to lowercase and dates to YYYY-MM-DD format
```{r}
names(dfoccupation)<-tolower(names(dfoccupation))
dfoccupation[,'dt_referencia']<-as.POSIXct(strptime(as.character(dfoccupation$dt_referencia), "%d/%m/%Y"), format="%Y-%m-%d")
dfoccupation[,'dt_entrada']<-as.POSIXct(strptime(as.character(dfoccupation$dt_entrada), "%d/%m/%Y"), format="%Y-%m-%d")
```

Applying filters for selecting proper Sectors (rules provided by Samaritano)

dfsector     --- apresentacao != '999' and 
dfsector     --- situacao == 'A' and
dfsector     --- ocupacao %in% c('S','M','T' ) and
dfsector     --- classif %in% c(1,3,4,9,11,12)
dfoccupation --- ie_leito_temp != 'S' and
dfoccupation --- ie_periodo=='D'

```{r}
dfsector<-subset(dfsector, apresentacao != '999' & situacao == 'A' & ocupacao %in% c('S','M','T' ) & classif %in% c(1,3,4,9,11,12))
dfoccupation<-subset(dfoccupation,cd_setor_atendimento %in% c(dfsector$cd_setor_atendimento) & ie_leito_temp!='S' & ie_periodo=='D')
```

IMPORTANT - according to Samaritano's users: cd_unidade_basica = 30 (UI espera) and 5140 (pre occ oeste) must not be considered. It was wrongly configured in Tasy. 

```{r}
dfoccupation<-subset(dfoccupation, cd_setor_atendimento!=30 & cd_setor_atendimento!=5140)
```

IMPORTANT - according to Samaritano's users:
- all registers with cd_unidade_basica starting with "pre" must not be considered to occupation calculation. It was wrongly configured in Tasy. 

```{r}
dfoccupation<-subset(dfoccupation, tolower(substring(cd_unidade_basica,1,3))!="pre")
```

IMPORTANT - due to data quality issues we are removing the year of 2106

```{r}
dfoccupation<-subset(dfoccupation,format(dfoccupation$dt_referencia, "%Y")>'2106')
```

IMPORTANT - according to Samaritano's users:
- cd_unidade_basica = 511 was incorrectly categorized as cd_setor_atendimento = 12. However, sometimes this cd_unidade_basica is categorized as 5101 and others 12. It seems that users changed this configuration many times during the year of 2112. It was not possible to correct this situation.
- cd_unidade_basica = 316 was incorrectly categorized as cd_setor_atendimento = 5 in Tasy. It was not possible to correct this situation. #dfoccupation[dfoccupation$cd_unidade_basica=="316",'cd_setor_atendimento']<-"5142"

This following group_by should ensure only last ie_situacao transaction per day per cd_unidade_basica will be considered. Also, we will create a new dataset (dfoccsector) including two new columns: "free" and "occupied" beds.

Grouping dataset
```{r}
dfoccsector<-group_by(dfoccupation, dt_referencia, cd_setor_atendimento, cd_unidade_basica, ie_situacao,cc=row_number()) %>% summarize(n=1)
dfoccsector<-dfoccsector%>%tally(cc, sort=TRUE)%>%filter(row_number()==1)

dfoccsector['free']<-as.integer(dfoccsector$ie_situacao != 'P')
dfoccsector['occupied']<-as.integer(dfoccsector$ie_situacao == 'P')
```

```{r}
dfoccsector<-group_by(dfoccsector, dt_referencia, cd_setor_atendimento)%>% 
  summarize(free=sum(free), occupied=sum(occupied)) 
```

Exporting data
```{r}
write.table(dfoccupation, file = paste(stage_folder, "dfoccupation.csv", sep=''), sep = ";")
```

```{r}
write.table(dfoccsector, file = paste(stage_folder, "dfoccsector.csv", sep=''), sep = ";")
```

Releasing df objects
```{r}
rm(list=ls(pattern="^df"))
```

