---
title: "MDL report for Margin KPI"
output: html_notebook
---

```{r}
data_folder <- paste(getwd(), '/../data/raw/', sep='')
stage_folder <- paste(getwd(), '/../data/stage/', sep='')
simulation_folder <- paste(getwd(), '/../data/simulation/', sep='')
script_folder <- paste(getwd(), '/../scripts/', sep='')
```

Importing libraries and functions
```{r}
library(dplyr)
library(grid)
library(gridExtra)
library(forecast)
source(paste(script_folder,"library.R",sep=''))
```

Reading IPCA data for deflation
```{r}
IPCA<-read.csv(paste(data_folder,"IPEA IPCA 1988-2017 indice geral.csv",sep=''), sep=',', header=T)
IPCA$year = as.numeric(substring(IPCA$date,1,4))
IPCA$month = as.numeric(substring(IPCA$date,6,7))
IPCA$month[which(IPCA$month %in% 11) - 1] = IPCA$month[which(IPCA$month %in% 11) - 1] * 10 # Get the right number for October
```

Reading Contribution Margin dataset
```{r}
dfmarginsector<-read.csv(paste(stage_folder,"dfmarginsector.csv",sep=''), sep = ";", header=T)
```

## Hospital Contribution Margin analysis

Medical Clinic	
Sectors: 3o Andar Norte (5), 5o Andar Norte (12), Térreo	(2), 6o andar	(11), 11o andar	(34), 12o andar	(183)

Surgery
Sectors: 02º Andar Lane (5147), 02º Andar Norte (13), 10º Andar (5057)

Emergency room
Sectors: Pronto Socorro Adulto (38) Pronto Socorro de Obstetrícia (47) Pronto Socorro Infantil(37) Pronto Socorro Observação (197)

UTI
Sectors: 	CTI-C (25), CTI Adulto - Ala Oeste	(5038), UTI Adulto	(29), UTI Neurológica	(5103), UTI Infantil	(27), UTI neonatal	(178), 

Semi- ICU	
Sectors: Semi Intensiva Adulto I	(179), Semi Intensiva Adulto II	(180), Semi Intensiva Infantil	(28), Semi Intensiva Neonatal (177)
		
Oncology	
Sectors: 8o andar	(5123), 9o andar	(32)
		
Maternity	
Sector: 4o andar	(14)

Hospital Contribution margin per month
```{r}
d<-group_by(dfmarginsector, dt_ref_fat)%>% summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
d[,'dt_ref_fat']<-as.POSIXct(d$dt_ref_fat, format="%Y-%m-%d")
head(d)
```

Describing dataset
```{r}
summary(d)
```

Surgery contribution margin 

dfmargin ie_tipo_atendimento == 1 (Internado)
dfmargin it_tipo_procedimento == 99 (Cirúrgico)

dfclinic_s<-subset(dfoccsector, cd_setor_atendimento==5147 | cd_setor_atendimento==13 | cd_setor_atendimento==5057)

```{r}
#e<-subset(dfmargin, ie_tipo_procedimento==99 & ie_tipo_atend==1)
e<-subset(dfmarginsector, cd_setor_atendimento==5147 | cd_setor_atendimento==13 | cd_setor_atendimento==5057)

e<-group_by(e, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
e[,'dt_ref_fat']<-as.POSIXct(e$dt_ref_fat, format="%Y-%m-%d")
```

Describing Surgery dataset
```{r}
summary(e)
```

Emergency room  contribution margin 

dfmargin  ie_tipo_atend == 3 (Pronto Socorro)

Sectors: Pronto Socorro Adulto (38) Pronto Socorro de Obstetrícia (47) Pronto Socorro Infantil(37) Pronto Socorro Observação (197)

```{r}
#f<-filter(dfmargin,ie_tipo_atend==3)
f<-subset(dfmarginsector, cd_setor_atendimento==38 | cd_setor_atendimento==47 | cd_setor_atendimento==37 | cd_setor_atendimento==197)

f<-group_by(f, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
f[,'dt_ref_fat']<-as.POSIXct(f$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(f)
```

UTI contribution margin 

Sectors: 	CTI-C (25), CTI Adulto - Ala Oeste	(5038), UTI Adulto	(29), UTI Neurológica	(5103), UTI Infantil	(27), UTI neonatal	(178), 

```{r}
g<-subset(dfmarginsector, cd_setor_atendimento==25 | cd_setor_atendimento==5038 | cd_setor_atendimento==29 | cd_setor_atendimento==5103 | cd_setor_atendimento==27 | cd_setor_atendimento==178)

g<-group_by(g, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
g[,'dt_ref_fat']<-as.POSIXct(g$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(g)
```

Oncology	
Sectors: 8o andar	(5123), 9o andar	(32)

```{r}
h<-subset(dfmarginsector, cd_setor_atendimento==5123 | cd_setor_atendimento==32)

h<-group_by(h, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
h[,'dt_ref_fat']<-as.POSIXct(h$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(h)
```

Semi- ICU	
Sectors: Semi Intensiva Adulto I	(179), Semi Intensiva Adulto II	(180), Semi Intensiva Infantil	(28), Semi Intensiva Neonatal (177)


```{r}
h1<-subset(dfmarginsector, cd_setor_atendimento==179 | cd_setor_atendimento==180 | cd_setor_atendimento==28 | cd_setor_atendimento==177)

h1<-group_by(h1, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
h1[,'dt_ref_fat']<-as.POSIXct(h1$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(h1)
```

Medical Clinic	
Sectors: 3o Andar Norte (5), 5o Andar Norte (12), Térreo	(2), 6o andar	(11), 11o andar	(34), 12o andar	(183)

```{r}
h2<-subset(dfmarginsector, cd_setor_atendimento==5 | cd_setor_atendimento==12 | cd_setor_atendimento==11 | cd_setor_atendimento==34 | cd_setor_atendimento==183)

h2<-group_by(h2, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
h2[,'dt_ref_fat']<-as.POSIXct(h2$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(h2)
```

Maternity	
Sector: 4o andar	(14)

```{r}
h3<-subset(dfmarginsector, cd_setor_atendimento==14)

h3<-group_by(h3, dt_ref_fat)%>%
summarize(cost=sum(cost),revenue=sum(revenue),margin=sum(margin),count=n())
h3[,'dt_ref_fat']<-as.POSIXct(h3$dt_ref_fat, format="%Y-%m-%d")
```

Describing dataset
```{r}
summary(h3)
```

##Fitting model to Hospitalization Contribution Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(d)*.85) #end of training
hpred<-5 # prediction horizon

train = window(d$margin, start=BOT, end=EOT)
test = window(d$margin, start=EOT+1)
```


```{r}
plot(d$margin/1000, type='l')
```

```{r}
fit0<-arima(train, order=c(0,0,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```
```{r}
fit1<-Arima(train, order=c(1,0,0), include.drift=TRUE)
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```

```{r}
summary(fit1)
```
```{r}
fit2<-Arima(train, order=c(0,1,3), include.drift=TRUE)
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```

```{r}
summary(fit2)
```

```{r}
fit3 = Arima(train, order=c(1,1,0), seasonal=list(order=c(0,1,1), period=12))
fcast3<-forecast(fit3, h=hpred)
coef3 = accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```

```{r}
fit4<-ets(train,model="ZZZ")
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
```

```{r}
summary(fit4)
```

```{r}
fit5 = Arima(train, order=c(3,1,0), include.drift=TRUE)
fcast5<-forecast(fit5, h=hpred)
coef5 = accuracy(fcast5, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast5$residuals)
pacf(fcast5$residuals)
plot(fcast5$residuals)
par(old.par)
```
```{r}
summary(fit5)
```


Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
coef5
```

```{r}
AIC(fit1,fit2, fit3, fit4, fit5)
```
```{r}
BIC(fit1, fit2, fit3, fit4, fit5)
```

```{r}
dados<-d$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=6)
models[1,1]<-1
models[1,2]<-"ARIMA(013) drift"
models[1,3]<-"013"
models[1,5]<-TRUE
models[2,1]<-1
models[2,2]<-"ARIMA(110)(011)[12]"
models[2,3]<-"110"
models[2,4]<-"01112"
models[3,1]<-1
models[3,2]<-"ARIMA(310)"
models[3,3]<-"310"
models[4,1]<-2
models[4,2]<-"ETS(MAN)"
models[4,3]<-"MAN"
models[5,1]<-2
models[5,2]<-"ETS(AAN)"
models[5,3]<-"AAN"
models[5,4]<-TRUE
models[6,1]<-1
models[6,2]<-"ARIMA(011) drift"
models[6,3]<-"011"
models[6,5]<-TRUE

matrixname<-"Contribution Margin - Hospital "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain"))
```


Predicting new data for Hospital margin

```{r}
vpred<-4

dados<-d$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Hospital Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 1:nrow(samples)) {
  
  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(2,0,0), include.drift = TRUE)
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(0,1,0), seasonal=list(order=c(0,0,1), period=12))
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(200)","ARIMA(011)(001)[12]","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```

##Fitting model to Surgery Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(e)*.85) #end of training
hpred<-5 # prediction horizon

train = window(e$margin, start=BOT, end=EOT)
test = window(e$margin, start=EOT+1)
```

```{r}
plot(e$margin/1000000, main="Surgery Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,1,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(3,1,0), include.drift=TRUE)
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(0,1,3), include.drift=TRUE) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(1,1,0), seasonal=list(order=c(0,1,0), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-e$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(013) drift"
models[1,3]<-"013"
models[1,5]<-TRUE
models[2,1]<-1
models[2,2]<-"ARIMA(110)(010)12"
models[2,3]<-"110"
models[2,4]<-"01012"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(310) drift"
models[4,3]<-"310"
models[4,5]<-TRUE
models[5,1]<-1
models[5,2]<-"ARIMA(011) drift"
models[5,3]<-"011"
models[5,5]<-TRUE

matrixname<-"Contribution Margin - Surgery "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for Surgery margin

```{r}
vpred<-4

dados<-e$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Surgery Contribution Margin', type = 'l', ylab="Margin (10^-6)", ylim=range(0.5,2.5), xlab="Months")

for (i in 1:nrow(samples)) {
  
  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(0,1,1), include.drift=TRUE)
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(0,1,3), include.drift=TRUE)
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(011) drift","ARIMA(013) drift","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```

##Fitting model to UTI Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(g)*.85) #end of training
hpred<-5 # prediction horizon

train = window(g$margin, start=BOT, end=EOT)
test = window(g$margin, start=EOT+1)
```

```{r}
plot(g$margin/1000000, main="UTI Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,1,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(2,1,0), include.drift=TRUE)
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(0,1,1), include.drift=TRUE) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(0,1,1), seasonal=list(order=c(0,1,0), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-g$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(011) drift"
models[1,3]<-"011"
models[1,5]<-TRUE
models[2,1]<-1
models[2,2]<-"ARIMA(011)(010)12"
models[2,3]<-"011"
models[2,4]<-"01012"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(210) drift"
models[4,3]<-"210"
models[4,5]<-TRUE
models[5,1]<-1
models[5,2]<-"AutoARIMA(011)"
models[5,3]<-"011"


matrixname<-"Contribution Margin - UTI "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for UTI margin

```{r}
vpred<-4

dados<-g$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - UTI Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 1:nrow(samples)) {
  
  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(0,1,1), include.drift=TRUE)
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(2,1,0), include.drift=TRUE)
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(011) drift","ARIMA(210) drift","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```
##Fitting model to Oncology Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(h)*.85) #end of training
hpred<-5 # prediction horizon

train = window(h$margin, start=BOT, end=EOT)
test = window(h$margin, start=EOT+1)
```

```{r}
plot(h$margin/1000000, main="Oncology Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,1,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(1,1,0), include.drift=TRUE)
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(0,1,1), include.drift=TRUE) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(0,1,1), seasonal=list(order=c(0,1,0), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-h$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(011) drift"
models[1,3]<-"011"
models[1,5]<-TRUE
models[2,1]<-1
models[2,2]<-"ARIMA(011)(010)12"
models[2,3]<-"011"
models[2,4]<-"01012"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(110) drift"
models[4,3]<-"110"
models[4,5]<-TRUE
models[5,1]<-1
models[5,2]<-"AutoARIMA(011)"
models[5,3]<-"011"


matrixname<-"Contribution Margin - Oncology "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for UTI margin

```{r}
vpred<-4

dados<-h$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Oncology Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 2:nrow(samples)) {

  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(0,1,1), include.drift=TRUE)
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(0,1,1), seasonal=list(order=c(0,1,0), period=12))
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(011) drift","ARIMA(011)(010)12","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```


##Fitting model to Semi-ICU Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(h1)*.85) #end of training
hpred<-5 # prediction horizon

train = window(h1$margin, start=BOT, end=EOT)
test = window(h1$margin, start=EOT+1)
```

```{r}
plot(h1$margin/1000000, main="Semi-ICU Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,1,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(1,1,0))
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(0,1,1)) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(0,1,1), seasonal=list(order=c(1,0,0), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-h1$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(011)"
models[1,3]<-"011"
models[2,1]<-1
models[2,2]<-"ARIMA(011)(100)12"
models[2,3]<-"011"
models[2,4]<-"10012"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(110)"
models[4,3]<-"110"
models[5,1]<-1
models[5,2]<-"AutoARIMA(011)"
models[5,3]<-"011"

matrixname<-"Contribution Margin - Semi-ICU "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for UTI margin

```{r}
vpred<-4

dados<-h1$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Semi-ICU Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 2:nrow(samples)) {

  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(0,1,1))
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(1,1,0))
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(011) drift","ARIMA(110)","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```

##Fitting model to Medical clinic Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(h2)*.85) #end of training
hpred<-5 # prediction horizon

train = window(h2$margin, start=BOT, end=EOT)
test = window(h2$margin, start=EOT+1)
```

```{r}
plot(h2$margin/1000000, main="Medical Clinic Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,1,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(1,1,0))
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(0,1,1)) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(1,0,0), seasonal=list(order=c(1,0,0), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-h2$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(011)"
models[1,3]<-"011"
models[2,1]<-1
models[2,2]<-"ARIMA(100)(100)12"
models[2,3]<-"100"
models[2,4]<-"10012"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(110)"
models[4,3]<-"110"
models[5,1]<-1
models[5,2]<-"AutoARIMA(100)"
models[5,3]<-"100"

matrixname<-"Contribution Margin - Medical Clinic "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for UTI margin

```{r}
vpred<-4

dados<-h2$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Medical Clinic Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 2:nrow(samples)) {

  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(1,0,0), seasonal=list(order=c(1,0,0), period=12))
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(1,0,0))
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(100)(100)12","AutoARIMA(100)","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```

##Fitting model to Maternity Margin

Setting train/test datasets

```{r}
BOT<-1 # Begin of training
EOT<-as.integer(nrow(h3)*.85) #end of training
hpred<-5 # prediction horizon

train = window(h3$margin, start=BOT, end=EOT)
test = window(h3$margin, start=EOT+1)
```

```{r}
plot(h3$margin/1000000, main="Maternity Margin", type = 'l', ylab="Margin (10^-6)", xlab="Months")
```

```{r}
fit0<-arima(train, order=c(0,0,0))
fcast0<-forecast(fit0, h=hpred)
coef0<-accuracy(fcast0, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast0$residuals)
pacf(fcast0$residuals)
plot(fcast0$residuals)
par(old.par)
```

```{r}
summary(fit0)
```

```{r}
fit1<-Arima(train, order=c(1,0,0))
fcast1<-forecast(fit1, h=hpred)
coef1<-accuracy(fcast1, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast1$residuals)
pacf(fcast1$residuals)
plot(fcast1$residuals)
par(old.par)
```
```{r}
summary(fit1)
```

```{r}
fit2<-Arima(train, order=c(2,1,0)) 
fcast2<-forecast(fit2, h=hpred)
coef2<-accuracy(fcast2, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast2$residuals)
pacf(fcast2$residuals)
plot(fcast2$residuals)
par(old.par)
```
```{r}
summary(fit2)
```

```{r}
fit3<-auto.arima(train) 
fcast3<-forecast(fit3, h=hpred)
coef3<-accuracy(fcast3, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast3$residuals)
pacf(fcast3$residuals)
plot(fcast3$residuals)
par(old.par)
```

```{r}
summary(fit3)
```
```{r}
fit4<-Arima(train, order=c(0,0,1), seasonal=list(order=c(0,0,1), period=12)) 
fcast4<-forecast(fit4, h=hpred)
coef4<-accuracy(fcast4, test[1:hpred-1])
old.par <- par(mfrow=c(2, 2))
acf(fcast4$residuals)
pacf(fcast4$residuals)
plot(fcast4$residuals)
par(old.par)
```
```{r}
summary(fit4)
```

Comparing error metrics

```{r}
coef1
coef2
coef3
coef4
```
```{r}
AIC(fit1,fit2, fit3, fit4)
```
```{r}
BIC(fit1,fit2, fit3, fit4)
```

```{r}
# Coefficients matrix

dados<-h3$margin

samplepred<- c(2,3,4,6,8,12)

samples<-generatesamples(samplepred, dados)

samplesBOT<-samples[,1]
samplesEOT<-samples[,2]

models<-matrix(rep(0,30),ncol=5, nrow=5)
models[1,1]<-1
models[1,2]<-"ARIMA(100)"
models[1,3]<-"100"
models[2,1]<-1
models[2,2]<-"ARIMA(210)"
models[2,3]<-"210"
models[3,1]<-2
models[3,2]<-"ETS(ANN)"
models[3,3]<-"ANN"
models[4,1]<-1
models[4,2]<-"ARIMA(200)"
models[4,3]<-"200"
models[5,1]<-1
models[5,2]<-"AutoARIMA(001)"
models[5,3]<-"001"

matrixname<-"Contribution Margin - Maternity "

results<-checkperformance(dados, models, samplespred, samplesBOT, samplesEOT, matrixname)
```

```{r}
grid.newpage()
grid.table(results, cols=c("Method", "Average of RMSE", "% Gain" ))
```

Predicting new data for UTI margin

```{r}
vpred<-4

dados<-h3$margin/1000000

samples<-generatesamples(vpred, dados)

plot(dados, main='Comparing Models - Maternity Contribution Margin', type = 'l', ylab="Margin (10^-6)", xlab="Months")

for (i in 2:nrow(samples)) {

  BOT<-samples[i,1] 
  EOT<-samples[i,2]
    
  strain = window(dados, start=BOT, end=EOT)
  stest = window(dados, start=EOT+1)
  vpredposition<-(EOT+1)

  # fiiting models
  fitf1<-Arima(strain, order=c(1,0,0))
  fcastf1<-forecast(fitf1, h=vpred)
  
  fitf2<-Arima(strain, order=c(0,1,2))
  fcastf2<-forecast(fitf2, h=vpred)
  
  aa<-ts(strain, frequency=1)
  fitf3<-tslm(aa~trend)
  fcastf3<-forecast(fitf3,h=vpred)
  
  fcastf4<-rep(mean(strain),vpred)

  lines(vpredposition:(vpredposition+vpred-1), fcastf1$mean, col='blue')
  lines(vpredposition:(vpredposition+vpred-1), fcastf2$mean, col='green')
  lines(vpredposition:(vpredposition+vpred-1), fcastf3$mean, col='red')
  lines(vpredposition:(vpredposition+vpred-1), fcastf4, col='brown', lty=2)
  abline(a=0,b=0,v=EOT, col="black", lty=2)
}
legend("topleft", c("ARIMA(100)","ARIMA(012)","linear model", "Mean"), fill=c("blue","green","red", "brown"), horiz=TRUE, cex=.5)
```